Relationships in the extremes and their influence on competition and coexistence

### 2022-04-05 paper/Paper.Rnw
finished the README and slight mods to the acks
minor changed to paper

%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
% throughout the paper. Has to be done in the main text and sup mat and the figures themselves.
% Need also search for "invader" since sometimes we use that word. Also think of any other related terminology that needs to 
% be altered as part of the overall change. 
%***JASMIN: Done in text. Need to finish change in figures.

@@

%G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we use a laboratory chemostat
system of freshwater diatoms explored by \cite{Descamps_2005}.
Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
\emph{Cyclotella pseudostelligera} over a range of temperatures to parametrize a resource
competition model. The model is 
\begin{align}
\frac{dS}{dt} &= D(S_0 - S) - Q_1x_1 \frac{V_1 S}{K_1 + S} - Q_2 x_2 \frac{V_2 S}{K_2 + S} \\
\frac{dx_j}{dt} &= x_j \frac{V_j S}{K_j+S}-D x_j, 
\end{align}
for $j=1,2$, where $S$ is a extracellular silicate concentration in the chemostat, $x_1$ is the population density
of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
$D$ is the chemostat outflow rate, $S_0$ is concentration of silicates in the chemostat
inflow, the $Q_j$ are amounts of silicate per cell, the $V_j$ are maximum population growth rates, and
the $K_j$ are half-saturation constants for the dependence of growth rates on nutrient concentration.
The temperature-dependent parameters $Q_j$, $V_j$, and $K_j$ were obtained by 
\cite{Descamps_2005} through batch experiments and curve fitting or interpolation.
\cite{Descamps_2005} showed that permanent coexistence of the two species occurs when 
temperature fluctuates periodically, 
$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down coexistence into contributing mechanisms
according to (\ref{eq:previous_decomp}) for these same values of
$\theta_0$, $a$ and $P$. We further decomponse
coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
made it sufficient to decompose only the GWR of species 1, for this
%***DAN: Please search for "growth when rare" and replace with GWR (except for the first instance,
%where GWR is defined). If one defines an acronym, one then has to use it.
%***JASMIN: Done
example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
$i=2$, $j=1$. Further details of the model setup are in SI section \ref{SI-sect:diatom_model}
and specifics of how our decomposition extends to this empirical case are in 
SI section \ref{SI-sect:Empircal_app}.

%***DAN: I see many instances of IGR have been replaced by GWR, but there are some IGRs remaining, e.g., 
%on the figures.
%***JASMIN: Figure 3 and 4 changed

@@

%***DAN: I think ATA rescue and ATA exclusion are mislabeled in the legend of Fig. 4. I think they got switched. Please check and correct if I am right.
%***JASMIN: Fixed.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Steve Ellner, and Robin Snyder for helpful discussions. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, and the California Department of Fish and
Wildlife Delta Science Program. 
%***DAN: Jasmim, please add any other acknolwedgments here - anyone else you have discussed this with or anything like that.
%Be generous here, add people even if they only helped a tiny bit. If Steve consents to be coauthor, we take him
%off this list and add him to the author list.

@@

%***DAN: Jasmin, some changes to this figure are needed. Please make axis labels and numbers bigger. Please put 
%axis labels outside the axes instead of inside. Please make panel labels bigger. Please make coutour labels
%slightly bigger.
\begin{figure}
\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
indicating greater importance of ATAs and values greater than $1$ corresponding to either
ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
$\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
were remapped to 1 and -1, respectively, for readability.
For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.} \label{fig:fig6}
\end{figure}

>>>>

@@

%G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we use a laboratory chemostat
system of freshwater diatoms explored by \cite{Descamps_2005}.
Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
\emph{Cyclotella pseudostelligera} over a range of temperatures to parametrize a resource
competition model. The model is 
\begin{align}
\frac{dS}{dt} &= D(S_0 - S) - Q_1x_1 \frac{V_1 S}{K_1 + S} - Q_2 x_2 \frac{V_2 S}{K_2 + S} \\
\frac{dx_j}{dt} &= x_j \frac{V_j S}{K_j+S}-D x_j, 
\end{align}
for $j=1,2$, where $S$ is a extracellular silicate concentration in the chemostat, $x_1$ is the population density
of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
$D$ is the chemostat outflow rate, $S_0$ is concentration of silicates in the chemostat
inflow, the $Q_j$ are amounts of silicate per cell, the $V_j$ are maximum population growth rates, and
the $K_j$ are half-saturation constants for the dependence of growth rates on nutrient concentration.
The temperature-dependent parameters $Q_j$, $V_j$, and $K_j$ were obtained by 
\cite{Descamps_2005} through batch experiments and curve fitting or interpolation.
\cite{Descamps_2005} showed that permanent coexistence of the two species occurs when 
temperature fluctuates periodically, 
$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down coexistence into contributing mechanisms
according to (\ref{eq:previous_decomp}) for these same values of
$\theta_0$, $a$ and $P$. We further decomponse
coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
made it sufficient to decompose only the GWR of species 1, for this
example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
$i=2$, $j=1$. Further details of the model setup are in SI section \ref{SI-sect:diatom_model}
and specifics of how our decomposition extends to this empirical case are in 
SI section \ref{SI-sect:Empircal_app}.

@@

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Steve Ellner, and Robin Snyder for helpful discussions. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, and the California Department of Fish and
Wildlife Delta Science Program. 
%***DAN: Jasmim, please add any other acknolwedgments here - anyone else you have discussed this with or anything like that.
%Be generous here, add people even if they only helped a tiny bit. If Steve consents to be coauthor, we take him
%off this list and add him to the author list.
%***DAN: DAN: if we add Ellner to the author list take him out here, and also take him out of the acks in the README.

@@

%***DAN: Jasmin, some changes to this figure are needed. Please make axis labels and numbers bigger. Please put 
%axis labels outside the axes instead of inside. Please make panel labels bigger. Please make coutour labels
%slightly bigger.
%***DAN: Do we want to do anything with standard error for this fig?
\begin{figure}
\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
indicating greater importance of ATAs and values greater than $1$ corresponding to either
ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
$\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
were remapped to 1 and -1, respectively, for readability.
For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.} \label{fig:fig6}
\end{figure}
